---
title: "Stat378_HW5_Q1"
output: html_document
---
This is a test change. Testing for github commit..

Below we have a function that computes matrix multiplication using for-loops.
```{python}
import numpy as np

# This function take a matrix and a np array vector in the form of 
# np.array([x1,x2,...])
def dot_product(M,x):
    xt = x.reshape(M.shape[1],1)
    y = np.zeros((M.shape[0],1))
    for i in np.arange(0,M.shape[0]):
        for j in np.arange(0,M.shape[1]):
            prod = M[i][j]*xt[j][0]
            y[i]+=prod
    return y

# This is the main function that uses the above for-loop implementation of 
#dot product to perform element-wise matrix vector multiplication and addition. 
def matrix_comp1(x):
    M = np.random.normal(loc =0, scale =1, size=(10, 10))
    N = np.random.normal(loc =0, scale =1, size=(10, 10))
    G = np.random.normal(loc =0, scale =1, size=(10, 1))
    Mx = dot_product(M,x)
    Nx = dot_product(N,x)
    return Mx+Nx+G

# Check if the function works as it should be.  
x = np.ones(10)
y = matrix_comp1(x)
print(y)

```
Below we have a function that computes matrix multiplication using vectorized method.
```{python}
# Function A with vectorized method
import numpy as np
def matrix_comp2(x):
  xt = x.reshape(x.size,1)
  M = np.random.normal(loc =0, scale =1, size=(10, 10))
  N = np.random.normal(loc =0, scale =1, size=(10, 10))
  G = np.random.normal(loc =0, scale =1, size=(10, 1))
  Mx = np.matrix(np.dot(M,x)).T
  Nx = np.matrix(np.dot(N,x)).T
  F = np.matrix(Mx)+np.matrix(Nx)+np.matrix(G)
  return F
  
# Check if the function works as it should be.  
x = np.ones(10)
y = matrix_comp2(x)
print(y)
```

Define a function that takes n, the number of iterations, and return computing time. Since we have double for-loops, each one performs M iterations for MxM square matrices, n=10000 iterations means 100x100 matrices.


```{python}

import numpy as np
from timeit import default_timer as timer
import matplotlib.pyplot as plt

def dot_product(M,x):
    xt = x.reshape(M.shape[1],1)
    y = np.zeros((M.shape[0],1))
    for i in np.arange(0,M.shape[0]):
        for j in np.arange(0,M.shape[1]):
            prod = M[i][j]*xt[j][0]
            y[i]+=prod
    return y

def matrix_computation(M,N,G,x):
    Mx = dot_product(M,x)
    Nx = dot_product(N,x)
#     print("Mx looks like: ", Mx, "Nx looks like: ", Nx, "G looks like: ", G)
    return Mx+Nx+G

elapsed_time = []
def time_comp1(n):
    iteration = np.arange(1,n+1).astype(int)
    for iter_ in iteration:
        M = np.random.normal(loc =0, scale =1, size=(iter_, iter_))
        N = np.random.normal(loc =0, scale =1, size=(iter_, iter_))
        G = np.random.normal(loc =0, scale =1, size=(iter_, 1))
        x = np.ones(iter_)
        start = timer()
        y = matrix_computation(M,N,G,x)
        end = timer()
        elapsed_time.append(end-start)
    return elapsed_time

elapsed_time = time_comp1(100)
print(elapsed_time)

k = np.arange(1,101)
x_ax = k*k
fig = plt.scatter(x_ax,elapsed_time)
plt.xlabel('Number of iterations', fontsize=16)
plt.ylabel('Elapsed time(s)', fontsize=16)

```

Define a function that times a function that uses vectorized method. To compare computing time to n = 1,...,10000 iterations, we use 1x1,2x2,...,100x100 matrices.

```{python}
elapsed_time = []

def matrix_comp2(M,N,G,x):
    xt = x.reshape(x.size,1)
    Mx = np.matrix(np.dot(M,x)).T
    Nx = np.matrix(np.dot(N,x)).T
    F = np.matrix(Mx)+np.matrix(Nx)+np.matrix(G)
    return F

def time_comp2(n):
    iteration = np.arange(1,n+1).astype(int)
    for iter_ in iteration:
        M = np.random.normal(loc =0, scale =1, size=(iter_, iter_))
        N = np.random.normal(loc =0, scale =1, size=(iter_, iter_))
        G = np.random.normal(loc =0, scale =1, size=(iter_, 1))
        x = np.ones(iter_)
        start = timer()
        y = matrix_comp2(M,N,G,x)
        end = timer()
        elapsed_time.append(end-start)
    return elapsed_time
    
elapsed_time = time_comp2(100)
    
k = np.arange(1,101)
x_ax = k*k
fig = plt.scatter(x_ax,elapsed_time)
plt.xlabel('Equivalent number of iterations', fontsize=16)
plt.ylabel('Elapsed time(s)', fontsize=16)
```


Below I have R implementation of Function B using double for-loops.

```{r}
matrix_comp1 <- function(x){
  M = matrix(rnorm(10*10,mean=0,sd=1), 10, 10) 
  G = matrix(rnorm(10*10,mean=0,sd=1), 10, 1) 
  for(i in 1:nrow(M)) {
    for(j in 1:ncol(M)) {
      M[i,j] = M[i,j]^3
    }
  }
  H = M%*%x+G
  return(H)
}

x<-1:10
H = matrix_comp1(x)
#print(H)
```

Below I have R implementation of Function B using vectorized methods.

```{r}
matrix_comp2 <- function(x){
  M = matrix(rnorm(10*10,mean=0,sd=1), 10, 10) 
  G = matrix(rnorm(10*10,mean=0,sd=1), 10, 1) 
  M = M^3
  H = M%*%x+G
  return(H)
}

# Test code
x<-1:10
H = matrix_comp2(x)
#print(H)
```

The below function times computing time the for-loop method takes.

```{r}
library(tictoc)
tic.clearlog()
size<-c(1:100)
iter<-size*size

matrix_comp3 <- function(l){
  li <-list()
  for (i in 1:l){
    #print(paste("current i is: ",i))
    M = matrix(rnorm(i*i,mean=0,sd=1), i, i) 
    G = matrix(rnorm(i*i,mean=0,sd=1), i, 1) 
    start <-Sys.time()
    for(j in 1:nrow(M)) {
      for(k in 1:ncol(M)) {
        M[j,k] = M[j,k]^3
      }
    }
    end <- Sys.time()
    time_diff <- as.numeric(end-start)
    #log.lst <- tic.log(format = TRUE)
    #timings <- unlist(lapply(log.lst, function(x) x$toc - x$tic(i)))
    #res <- gsub( " sec elapsed", "", unlist(log.txt))
    li<-append(li,time_diff)
  }
  plot(iter,li,xlab = "Number of iterations",ylab = "Time elapsed(s)")
}

matrix_comp3(100)

```
The below function times computing time it takes with the vectorized method. Cubing each element in a MxM matrix takes M iteration (for columns) by M iteration (for rows) making it MxM iterations. So cubing a MxM matrix with a vectorized methos is "one shot" but its equivalent number of iterations is MxM.

```{r}
library(tictoc)
tic.clearlog()
size<-c(1:100)
iter<-size*size

matrix_comp3 <- function(l){
  li <-list()
  for (i in 1:l){
    #print(paste("current i is: ",i))
    M = matrix(rnorm(i*i,mean=0,sd=1), i, i) 
    G = matrix(rnorm(i*i,mean=0,sd=1), i, 1) 
    start <-Sys.time()
    M = M^3
    end <- Sys.time()
    time_diff <- as.numeric(end-start)
    #log.lst <- tic.log(format = TRUE)
    #timings <- unlist(lapply(log.lst, function(x) x$toc - x$tic(i)))
    #res <- gsub( " sec elapsed", "", unlist(log.txt))
    li<-append(li,time_diff)
  }
  plot(iter,li,xlab = "Equivalent number of iterations",ylab = "Time elapsed(s)")
}

matrix_comp3(100)

```

