---
title: "Stat378_HW5_Q1"
output: html_document
---
This is a test change. Testing for github commit..

Below we have a function that computes matrix multiplication using for-loops.
```{python}
import numpy as np

# This function take a matrix and a np array vector in the form of 
# np.array([x1,x2,...])
def dot_product(M,x):
    xt = x.reshape(M.shape[1],1)
    y = np.zeros((M.shape[0],1))
    for i in np.arange(0,M.shape[0]):
        for j in np.arange(0,M.shape[1]):
            prod = M[i][j]*xt[j][0]
            y[i]+=prod
    return y

# This is the main function that uses the above for-loop implementation of 
#dot product to perform element-wise matrix vector multiplication and addition. 
def matrix_comp1(x):
    M = np.random.normal(loc =0, scale =1, size=(10, 10))
    N = np.random.normal(loc =0, scale =1, size=(10, 10))
    G = np.random.normal(loc =0, scale =1, size=(10, 1))
    Mx = dot_product(M,x)
    Nx = dot_product(N,x)
    return Mx+Nx+G

# Check if the function works as it should be.  
x = np.ones(10)
y = matrix_comp1(x)
#print(y)

```
Below we have a function that computes matrix multiplication using vectorized method.
```{python}
# Function A with vectorized method
import numpy as np
def matrix_comp2(x):
  xt = x.reshape(x.size,1)
  M = np.random.normal(loc =0, scale =1, size=(10, 10))
  N = np.random.normal(loc =0, scale =1, size=(10, 10))
  G = np.random.normal(loc =0, scale =1, size=(10, 1))
  Mx = np.matrix(np.dot(M,x)).T
  Nx = np.matrix(np.dot(N,x)).T
  F = np.matrix(Mx)+np.matrix(Nx)+np.matrix(G)
  return F
  
# Check if the function works as it should be.  
x = np.ones(10)
y = matrix_comp2(x)
#print(y)
```

Define a function that takes n, the number of iterations, and return computing time for the for-loop method.

```{python}
from timeit import default_timer as timer
import matplotlib.pyplot as plt

def forloop_time(n): # The function takes n iteration
    start = timer()
    for iteration in np.arange(1,n):
        x = np.random.normal(loc =0, scale =1, size=(10, 1)) # Generate a randome 10x1 vector from standard normal distribution
        matrix_comp1(x)
    end = timer()
    elapsed_time = end-start
    return elapsed_time
```

Define another function that takes n, the number of iterations, and return computing time for the vectorized method.

```{python}
#import numpy as np
def vectorized_time(n): # The function takes n iteration
    start = timer()
    for iteration in np.arange(0,n):
        x = np.random.normal(loc =0, scale =1, size=(10, 1)) # Generate a randome 10x1 vector from standard normal distribution
        matrix_comp2(x)
    end = timer()
    elapsed_time = end-start
    return elapsed_time
```

Let's compare the computing times for these two methods. Instead of computing for all number of iterations from 1 to 10000, I'll create a list of evenly spaced number of iterations (every 100 iterations until we reach 10000 iterations). I will plot the computing times to show how fast vectorized method is compared to the for-loop implementtion for each number of iterations.

```{python}
import numpy as np
# Compute time it takes to finish n iterations of for-loop function and store it to plot later.
comp_time = []
num_iterations = np.arange(0,10001,100)
for i in num_iterations:
    comp_time.append(forloop_time(i))
# Compute time it takes to finish n iterations of the vectorized function and store it to plot later.
comp_time2 = []
num_iterations = np.arange(0,10001,100)
for i in num_iterations:
    comp_time2.append(vectorized_time(i))
    
fig, ax = plt.subplots()
ax.scatter(num_iterations,comp_time2,color='green',label='vectorized')
ax.scatter(num_iterations,comp_time, color='red',label='for-loop')
plt.xlabel("Number of iterations")
plt.ylabel("Computing time (s)")
ax.legend()

```

Below I have R implementation of Function B using double for-loops.

```{r}
matrix_comp3 <- function(x){
  M = matrix(rnorm(10*10,mean=0,sd=1), 10, 10) 
  G = matrix(rnorm(10*10,mean=0,sd=1), 10, 1) 
  for(i in 1:nrow(M)) {
    for(j in 1:ncol(M)) {
      M[i,j] = M[i,j]^3
    }
  }
  H = M%*%x+G
  return(H)
}

x<-1:10
H = matrix_comp3(x)
#print(H)
```

Below I have R implementation of Function B using vectorized methods.

```{r}
matrix_comp4 <- function(x){
  M = matrix(rnorm(10*10,mean=0,sd=1), 10, 10) 
  G = matrix(rnorm(10*10,mean=0,sd=1), 10, 1) 
  M = M^3
  H = M%*%x+G
  return(H)
}

# Test code
x<-1:10
H = matrix_comp4(x)
#print(H)
```

The below function times computing time the for-loop method takes for Function B.

```{r}
library(tictoc)
tic.clearlog()
size<-c(1:100)
iter<-size*size

comp_time3 <- function(n){
  start <-Sys.time()
  for(i in 1:n) {
      x = matrix(rnorm(10*10,mean=0,sd=1), 10, 1) 
      matrix_comp3(x)
      }
  end <- Sys.time()
  time_diff <- as.numeric(end-start)
  return(time_diff)
  }

#comp_time3(500)
```
The below function times computing time the vectorized method takes for Function B.

```{r}
tic.clearlog()

comp_time4 <- function(n){
  start <-Sys.time()
  for(i in 1:n) {
      x = matrix(rnorm(10*10,mean=0,sd=1), 10, 1) 
      matrix_comp4(x)
      }
  end <- Sys.time()
  time_diff <- as.numeric(end-start)
  return(time_diff)
  }
```
These two functions time the computing time it takes for the for-loop methos and the vectorized method to perform n iterations. I created a list of equally spaced numbers from 1 to 10001 and stored the computing times to plot them in the next step.

```{r}
li <-list()
li2 <- list()
c<-seq(from = 1, to = 10001, by = 100)

for(i in c) {
  time3<-comp_time3(i)
  li<-append(li,time3)
}

#print(c)
for(i in c) {
  time4<-comp_time4(i)
  li2<-append(li2,time4)
}
```

Below is the plot for computing time it takes for the n iterations of for-loop implementaion of function B.
```{r}
plot(c,li,xlab = "Number of iterations",ylab = "Computing time (s)",main  ="For loop method")
```
Below is the plot for computing time it takes for the n iterations of the vectorized method for function B.

```{r}
plot(c,li2,xlab = "Number of iterations",ylab = "Computing time (s)",main  ="Vectorized method")
```




Student 2 Question 1 Function A

Q1ii) Vectorised method
```{r}
#initialising the matrices and vectors
entries_m = rnorm(100, mean=0, sd=1)
entries_n = rnorm(100, mean=0, sd=1)
entries_x = rnorm(10, mean=0, sd=1)
entries_g = rnorm(10, mean=0, sd=1)

m = matrix(entries_m, nrow=10)
n = matrix(entries_n, nrow=10)
x = matrix(entries_x, nrow=10)
g = matrix(entries_g, nrow=10)
```

```{r}
#vectorised approach
fun_a1 <- function(){
start.time <- Sys.time()
m%*%x + n%*%x + g 
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
}

fun_a1()
```

Q1i) For Loop method
```{r}
#Element by element approach
fun_a2 <- function(){
  
final_matrix <- c()
final_m1 <- c()
final_m2 <- c()

for(i in 1:10){
  m1 <- c()
  m2 <- c()
  
  for(j in 1:10){
   
    m1 <- c(m1, (m[i,j] * x[j]))
    m1 <- sum(m1)
    m2 <- c(m2, (n[i,j] * x[j]))
    m2 <- sum(m2)
    
    
  }
  
  final_m1 <- c(final_m1, m1)
  final_m2 <- c(final_m2, m2)
  
  final_matrix <- c(final_matrix, final_m1[i] + final_m2[i] + g[i])
  
  
}

}

fun_a2()
```

Q1(iii) Calculating and plotting time for "for loop" approach

```{r}
#time function
time <- function(){
  
  time.taken <- c()
  n <- 0
  
  while(n < 10000){
    
    start.time <- Sys.time() #start timing
    fun_a2()
    end.time <- Sys.time() #end timing
    
   
    time.taken <- c(time.taken, end.time - start.time)
    
    n <- n+1
  }
  

  
  return(time.taken*1000)
}

plot(time(), ylab="time in ms", xlab ="n") #plot


```

Q1(iv) Calculating and plotting time for vectorised approach

```{r}
#time function
time <- function(){
  
  time.taken <- c()
  n <- 0
  
  while(n < 10000){
    
    start.time <- Sys.time() #start time
    fun_a1()
    end.time <- Sys.time() #end time
    
   
    time.taken <- c(time.taken, end.time - start.time)
    
    n <- n+1
  }
  

  
  return(time.taken*1000)
}

plot(time(), ylab="time in ms", xlab ="n") #plot


```




Student 2 Question 1 Function B

```{r}
library(reticulate)
```

```{python}
#initialisation
import numpy as np

mu, sigma = 0, 1 # mean and standard deviation
s = np.random.normal(mu, sigma, 100)
v = np.random.normal(mu, sigma, 10)

m = np.array(np.matrix(s).reshape(10,10))
x = np.array([v]).transpose()
g = np.array([v]).transpose()

```

Q1ii) Vectorised method 

```{python}

def func_1():

  mat1 = np.dot(m,m)
  mat2 = np.dot(mat1,m)

  vectorised_result = np.dot(mat2,x) + 3*g
  
func_1()

```

Q1i) For loop method

```{python}
def func_2():
  res = [[0 for x in range(10)] for y in range(10)]  #10x10 empty matrix
  res2 = [[0 for x in range(10)] for y in range(10)]  
  res3 = np.array([[0 for x in range(10)]]).transpose() #10x1 empty vector

  # explicit for loops 
  for i in range(len(m)): 
    for j in range(len(m[0])): 
      for k in range(len(m)): 
  
        # resulted matrix 
        res[i][j] += m[i][k] * m[k][j] #m^2
  
  for i in range(len(res)): 
    for j in range(len(m[0])): 
      for k in range(len(m)): 
  
      # resulted matrix 
        res2[i][j] += res[i][k] * m[k][j] #m^3
            
  for i in range(len(res2)):
    for j in range(len(x)):
    
      res3[i] = res3[i] + (res2[i][j] * x[j]) #m^3 %*% x
    
  for i in range(len(res3)):
    res3[i] = res3[i] + 3*g[i] #m^3 %*% x + 3g
    


```

Q1iii) Timing and plotting for loop method

```{python}
import timeit
import matplotlib.pyplot as plt 

def time4():

  n = 0
  time = []
  while(n < 10000):
    tic = timeit.default_timer()
   
    func_2()
   
    toc = timeit.default_timer()

    time.append(((toc - tic)*1000))
    n += 1
    
  return(np.array(time)*1000)


x_axis = list(range(10000))

plt.plot(x_axis,time4())

# naming the x axis 
plt.xlabel('iterations') 
# naming the y axis 
plt.ylabel('time taken in ms') 
  
# giving a title to my graph 
plt.title('For loop method') 
  
# function to show the plot 
plt.show() 
```

Q1iv) Timing and plotting vectorised method

```{python}
import timeit
import matplotlib.pyplot as plt 

def time3():

  n = 0
  time = []
  while(n < 10000):
    tic = timeit.default_timer()
   
    func_1()
   
    toc = timeit.default_timer()

    time.append(((toc - tic)*1000))
    n += 1
    
  return(np.array(time)*1000)


x_axis = list(range(10000))

plt.plot(x_axis,time3())

# naming the x axis 
plt.xlabel('iterations') 
# naming the y axis 
plt.ylabel('time taken in ms') 
  
# giving a title to my graph 
plt.title('Vectorised method') 
  
# function to show the plot 
plt.show() 
```
